Milestone 8: To Turn In

### 2020-04-28 m8_turn_in.Rmd
initial commit with new repo

---
title: "Final Extension"
author: "Maria Burzillo"
date: "3/28/2020"
output: 
  pdf_document:
    extra_dependencies: ["rotating"]
bibliography: trounstine_bib.bib
#biblio-style: "apalike"
link_citations: true
header_includes:
  - \usepackage{float}
  - \usepackage{dcolumn}
  - \usepackage{rotating}

@@

# Extension

```{r setup, include = FALSE}

# import both datasets
library(haven)
rp <- read_dta("racial_polarization_winners.dta")
fin_seg<- read_dta("fin_seg.dta")

# calculate number of original observations
obs_fin_seg <- nrow(fin_seg)
obs_rp <- nrow(rp)

One concern with the original analysis in @trounstine_segregation_2016 is the large amount of missing data values in her original datasets. Because both R and Stata drop any observations with missing values for any of the variables used in a regression, this can exlcude a large portion of the data from the analysis and potentially bias the results if the data is not missing completely at random. A large amount of data is missing in both of the main datasets used in the analyses of @trounstine_segregation_2016. The main model specification using the racial polarization dataset exludes `r (obs_rp - 91)` of the original  `r obs_rp` observations, or  %`r round(100*(obs_rp - 91)/obs_rp,2)` of the data. Additionally, the main model specification using the financial segregation data set excludes `r obs_fin_seg - 13742` of the original `r obs_fin_seg` observations, or %`r round(100*(obs_fin_seg - 13742)/ obs_fin_seg,2)` of observations from the original dataset.

In an attempt to better deal with the problem of missing data in @trounstine_segregation_2016, I impute missing values in the original data using the mice package in R, which generates multivariate imputations using chained equations [@mice]. While there are a variety of different imputation methods that could have been employed, multiple imputation (such as the multiple multivariate imputations generated by mice) is desireable because instead of inputting a single value such as the mean for missing values, it instead uses the distribution of the available data to estimate multiple potential values for the missing data. As a result, multiple imputation helps to account for the uncertainty inherent in the imputation process and allows for the calculation of standard errors around estimaotrs. As a result, multiple imputation allows the researcher to more accurately assess the of uncertainty in the analysis in general.

Before performing the multiple imputations on the datasets, I first examined the missing data for patterns. To better understand any potential patterns in missing data, I plotted the pattern of missingness for only and created a histogram showing the frequency of missing values for those variables with missing values in figure X1 for the racial polarization data and figure X2 for the financial segregation data. Looking at figure X1, the histogram shows that the variables for the largest vote split along racial lines and the variable indicating whether or not the election had candidates of more than one race had by far the largest percentages of missing data, missing %`r round(sum(is.na(rp$biggestsplit))/nrow(rp)*100,2)` and  %`r round(sum(is.na(rp$biracial))/nrow(rp)*100,2)`, respectively. It is important to note that the variable for the largest vote split along racial lines is the dependent variable in the regresison analysis, and thus, we are missing a large percentage of this key variable. From the plot on the right, we can see that approximately 42% of observations are complete (have no missing values). There seem to be a correspondence between missing a value for biggest split and missing biracial. There also seem to be about 2% of values for wich most of the variables are missing. However, most observations are not missing more than 2-3 values.

### need to fix numbers below because this was from the whole fin_seg and not from the one filtered via the regression coefficients

In figure X2, we can similarly observe the trends for the financial segregation dataset. In this dataset, there is a very high proportion of missing variables for a number of variables. For example, there are a `r sum(is.na(fin_seg$H_citytract_NHW_i))` missing values for the segregation index, which is the main independent variable, or %`r round(sum(is.na(fin_seg$H_citytract_NHW_i))/nrow(fin_seg)*100,2)` of the data.

Then, I use the imputed datasets to re-estimate the original models, pooling the results to final pooled regression coefficients and parameters.

for all of the variables in the racial polarization data used in the analysis and for the main independent variable, the segregation index, used in the analysis with the financial segregation dataset. More of the data could not be imputed from the financial segregation dataset due to computing and time constraints for this project.

@@

>>>>

---
title: "Milestone 8: To Turn In"
author: "Maria Burzillo"
date: "3/28/2020"
output: 
  pdf_document:
    extra_dependencies: ["rotating"]
bibliography: trounstine_bib.bib
#biblio-style: "apalike"
link_citations: true
header_includes:
  - \usepackage{float}
  - \usepackage{dcolumn}
  - \usepackage{rotating}

@@

# Abstract

This is an extension of Jessica Trounstine's "Segregation and Inequality in Public Goods" (2016). I was able to replicate the main results of Trounstine's paper in R to suggest that racial segregation contributes to political polarization and decreased spending on public goods. Additionally, I extend the analysis by imputing missing data and rerunning Trounstine's original model as a robustness check.

@@

stargazer(imp_1_rp_sum_table,
          header = "Racial Divide with Multigroup Segregation Index",
          title = "\\textbf{Racial Polarization in Segregated Cities}")
